####  HW ASSIGNMENT 1 - NAGULAN NATHAN - BTMA431 - 30064918 #####

```{r}
library(quantmod)
library(ggplot2)
library(lubridate)
library(dplyr)
```

```{r}
setwd("C:/Users/nagul/Documents/Working Directory")
```

```{r}
load('HW1_data.rData')
```


```{r}
longrunROI <- numeric(7)
chart.vector <- c("BTC.charts","DASH.charts","ETH.charts","LTC.charts","PPC.charts","XLM.charts","XRP.charts")
i <- 1
for (j in chart.vector){
    eval(parse(text=paste0("longrunROI[", i,"] <- (",j,"$close[",j,"$date == '2021-09-01'] - ",j,"$close[",j,"$date ==          '2016-01-10'])/",j,"$close[",j,"$date == '2016-01-10']")))
    i <- i+1
} 
```

```{r}
ROI.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),ROI = longrunROI) #creating dataframe of longrun ROI of each cryptocurrency
ROI.data
```

```{r}
3836.77 * 100 #ROUND TO 383700 - THIS IS THE ANSWER FOR 1A
```

1a) The highest long-run ROI belongs to the Ethereum cryptocurrency. 3836.77 * 100 = 383700

```{r}
#filtering data to only include the relevant dates
BTC.charts2 <- BTC.charts[BTC.charts$date >= '2016-01-10' & BTC.charts$date <= '2021-09-01',]
DASH.charts2 <- DASH.charts[DASH.charts$date >= '2016-01-10' & DASH.charts$date <= '2021-09-01',]
ETH.charts2 <- ETH.charts[ETH.charts$date >= '2016-01-10' & ETH.charts$date <= '2021-09-01',]
LTC.charts2 <- LTC.charts[LTC.charts$date >= '2016-01-10' & LTC.charts$date <= '2021-09-01',]
PPC.charts2 <- PPC.charts[PPC.charts$date >= '2016-01-10' & PPC.charts$date <= '2021-09-01',]
XLM.charts2 <- XLM.charts[XLM.charts$date >= '2016-01-10' & XLM.charts$date <= '2021-09-01',]
XRP.charts2 <- XRP.charts[XRP.charts$date >= '2016-01-10' & XRP.charts$date <= '2021-09-01',]
```

```{r}
#using for loop and vectorization to get mean daily return for each currency and formatting it in data frame.
#we find out that ETH has the best mean daily return
chart.vector2 <- c("BTC.charts2","DASH.charts2","ETH.charts2","LTC.charts2","PPC.charts2","XLM.charts2","XRP.charts2")

i <- 1

for (j in chart.vector2){
    eval(parse(text=paste0("dailyReturn",i," <- (",j,"$close[2:nrow(",j,")]-",j,"$close[1:nrow(",j,")-1])/",j,"$close[1:nrow(",j,")-1]")))
  i <- i+1
} 

meanDailyReturn.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),MeanDailyReturn=c(mean(dailyReturn1),mean(dailyReturn2),mean(dailyReturn3),mean(dailyReturn4),mean(dailyReturn5),mean(dailyReturn6),mean(dailyReturn7)))
meanDailyReturn.data
```
1b)The highest mean daily return is for Ethereum. 0.00577*100 gives us 0.58.

```{r}
#using sd function on each of the vectors of daily returns and then formatting it in data frame shows us that BTC 
#has the least standard deviation
std.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),standarDeviation = c(sd(dailyReturn1),sd(dailyReturn2),sd(dailyReturn3),sd(dailyReturn4),sd(dailyReturn5),sd(dailyReturn6),sd(dailyReturn7)))
std.data
```

1c) BTC has the least standard deviation. 0.0399769 rounded to 4.00 for the D2L quiz

QUESTION 2

PORTFOLIO 1 (100%)

5K BTC

PORTFOLIO 2 (50% of each)

2.5K XRP
2.5K LTC

PORTFOLIO 3 (25% OF EACH)

1.25K ETH
1.25K DASH
1.25K PPC
1.25K XLM

```{r}
#PORTFOLIO 1 VALUE AS OF 2021 SEPT
numBTC=5000/BTC.charts2[1,5]
pf1Val=BTC.charts2[2062,5]*numBTC
pf1Val
```
PORTFOLIO 1 WOULD BE WORTH $545,178 AS OF Sept 1 2021

```{r}
#PORTFOLIO 2 VALUE AS OF 2021 SEPT
numXRP=2500/XRP.charts2[1,5]

numLTC=2500/LTC.charts2[1,5]

pf2val=(LTC.charts2[2062,5]*numLTC)+(XRP.charts2[2062,5]*numXRP)
pf2val
```
PORTFOLIO 2 WOULD BE WORTH $646,257 AS OF SEPT 1 2021

```{r}
#PORTFOLIO 3 VALUE AS OF 2021 SEPT
numETH=1250/ETH.charts2[1,5]
numDASH=1250/DASH.charts2[1,5]
numPPC=1250/PPC.charts2[1,5]
numXLM=1250/XLM.charts2[1,5]

pf3val=(ETH.charts2[2062,5]*numETH)+(DASH.charts2[2062,5]*numDASH)+(PPC.charts2[2062,5]*numPPC)+(XLM.charts2[2062,5]*numXLM)
pf3val

```
PORTFOLIO 3 WOULD BE WORTH $5,141,220 AS OF SEPT 1 2021 (FOR D2L , ROUND TO $5141000)


2a) Portfolio 1 = $545,178
    Portfolio 2 = $646,257
    Portfolio 3 = $5,141,220
    

```{r}
#VALUE OF EACH PORTFOLIO IN OUR TIME PERIOD OF INTEREST
pf1runval <- data.frame(date=BTC.charts2$date[1:2062],runningValue= numBTC*BTC.charts2$close[1:2062])

pf2runval <- data.frame(date=XRP.charts2$date[1:2062],runningValue = ((XRP.charts2$close[1:2062]*numXRP)+(LTC.charts2$close[1:2062]*numLTC)) )

pf3runval <- data.frame(date=ETH.charts2$date[1:2062], runningValue = (ETH.charts2$close[1:2062]*numETH)+(DASH.charts2$close[1:2062]*numDASH)+(PPC.charts2$close[1:2062]*numPPC)+(XLM.charts2$close[1:2062]*numXLM))

```

```{r}
#COMBINED DATAFRAME OF ALL PORTFOLIO VALUES
pfValues <- data.frame(Date = BTC.charts2$date, Portfolio1 = pf1runval$runningValue, Portfolio2 = pf2runval$runningValue, Portfolio3 = pf3runval$runningValue)
```

```{r}
head(pfValues,3)
```

```{r}
#PLOT OF ALL 3  PORTFOLIO VALUES
ggplot(data=pfValues,aes(x=Date))+geom_line(aes(y=Portfolio1,color = "Portfolio1"))+geom_line(aes(y=Portfolio2,color = "Portfolio2"))+geom_line(aes(y=Portfolio3,color = "Portfolio3"))+labs(y="Value ($)",x="Date")+ggtitle("Value of Portfolios")+theme(plot.title=element_text(hjust=0.5))+ylim(0,5200000)
```
```{r}
#THE MAXIMUM VALUES ATTAINED BY EACH PORTFOLIO QUESTION 2B
cat("The maximum dollar values of Portfolio 1,2 and 3 in order are",max(pfValues$Portfolio1),"and",max(pfValues$Portfolio2),"and",max(pfValues$Portfolio3))

```
```{r}
pf2dailyReturn <- c(NA,(pf2runval$runningValue[2:2062]-pf2runval$runningValue[1:2061])/pf2runval$runningValue[1:2061])
pf3dailyReturn <- c(NA,(pf3runval$runningValue[2:2062]-pf3runval$runningValue[1:2061])/pf3runval$runningValue[1:2061])
pf1dailyReturn <- c(NA,(pf1runval$runningValue[2:2062]-pf1runval$runningValue[1:2061])/pf1runval$runningValue[1:2061])
```

```{r}
#2C SCATTERPLOT 1
qplot(x=na.omit(pf2dailyReturn),y=dailyReturn1)+labs(x='Portfolio 2 Daily Return',y='BTC Daily Return')+ggtitle("Scatterplot of BTC Daily Returns vs Portfolio 2")
```

```{r}
#2C SCATTERPLOT 2
qplot(x=na.omit(pf3dailyReturn),y=dailyReturn1)+labs(x='Portfolio 3 Daily Return',y='BTC Daily Return')+ggtitle("Scatterplot of BTC Daily Returns vs Portfolio 3")
```

```{r}
cor(na.omit(pf2dailyReturn),dailyReturn1)
cor(na.omit(pf3dailyReturn),dailyReturn1)
```
2c)The correlation coefficient for portfolio 2 and BTC is 0.4921938 (0.49)
   The correlation coefficient for portfolio 3 and BTC is 0.6349575 (0.63)

```{r}
#THIS IS TO ANSWER THE D2L QUESTION
cor(na.omit(pf1dailyReturn),na.omit(pf3dailyReturn))
```

```{r}
#adding daily returns to the running value dataframe
pf1runval$dailyReturnpf1 <- pf1dailyReturn
pf2runval$dailyReturnpf2 <- pf2dailyReturn
pf3runval$dailyReturnpf3 <- pf3dailyReturn
```

```{r}
#ADDING BITCOIN DATA INTO THE PORTFOLIO DATAFRAME FOR EASY ANALYSIS
pf2withBTC <- data.frame(date=pf2runval$date,Portfolio2value=pf2runval$runningValue,Portfolio2DailyReturn=pf2runval$dailyReturnpf2,BTCDailyReturn=c(NA,dailyReturn1))
head(pf2withBTC,3)
```

```{r}
#ADDING BITCOIN DATA INTO THE PORTFOLIO DATAFRAME FOR EASY ANALYSIS
pf3withBTC <- data.frame(date=pf3runval$date,Portfolio3value=pf3runval$runningValue,Portfolio3DailyReturn=pf3runval$dailyReturnpf3,BTCDailyReturn=c(NA,dailyReturn1))
head(pf3withBTC,3)
```


```{r}
#FILTERING FOR DAYS WHERE BTC HAS NEGATIVE OR POSITIVE RETURNS
pf2withBTCplus <- filter(pf2withBTC,BTCDailyReturn > 0)
pf2withBTCneg <- filter(pf2withBTC,BTCDailyReturn < 0)
pf3withBTCplus <- filter(pf3withBTC,BTCDailyReturn > 0)
pf3withBTCneg <- filter(pf3withBTC,BTCDailyReturn < 0)
```

```{r}
#FINDING MEAN RETURNS OF THE PORTFOLIO
mean(pf2withBTCplus$Portfolio2DailyReturn)
mean(pf2withBTCneg$Portfolio2DailyReturn)
mean(pf3withBTCplus$Portfolio3DailyReturn)
mean(pf3withBTCneg$Portfolio3DailyReturn)
```

2d)
The mean daily return for Portfolio 2 for days when Bitcoin daily return was positive is 0.02485 
The mean daily return for Portfolio 2 for days when Bitcoin daily return was negative is -0.02116
The mean daily return for Portfolio 3 for days when Bitcoin daily return was positive is 0.02470
THe mean daily return for Portfolio 3 for days when Bitcoin daily return was negative is -0.0200

Neither portfolio 'dominated' the other for when BTC returns went up or went down as they both had very similar mean returns.
Portfolio 2 did have slightly higher returns and slightly higher negative returns than Portfolio 3 when BTC went up and down respectively, but it is a minor difference. Meaning it did better on average when BTC went up and did worse on average when BTC went down.


QUESTION 3 ROBUSTNESS CHECK BELOW (THE ASSIGNMENT IS DUPLICATED USING AVERAGE, AND IF WE GO THE BOTTOM WE SEE THE ONLY DIFFERENCE IN RESPONSE IS FROM 2D.)


```{r}
BTC.charts$avg <- (BTC.charts$high+BTC.charts$low)/2
DASH.charts$avg <- (DASH.charts$high+DASH.charts$low)/2
ETH.charts$avg <- (ETH.charts$high+ETH.charts$low)/2
LTC.charts$avg <- (LTC.charts$high+LTC.charts$low)/2
XLM.charts$avg <- (XLM.charts$high+XLM.charts$low)/2
XRP.charts$avg <- (XRP.charts$high+XRP.charts$low)/2
PPC.charts$avg <- (PPC.charts$high+PPC.charts$low)/2

```


```{r}
longrunROI <- numeric(7)
chart.vector <- c("BTC.charts","DASH.charts","ETH.charts","LTC.charts","PPC.charts","XLM.charts","XRP.charts")
i <- 1
for (j in chart.vector){
    eval(parse(text=paste0("longrunROI[", i,"] <- (",j,"$avg[",j,"$date == '2021-09-01'] - ",j,"$avg[",j,"$date ==          '2016-01-10'])/",j,"$avg[",j,"$date == '2016-01-10']")))
    i <- i+1
} 
```

```{r}
#NO CHANGE IN RESPONSE NOW THAT WE ARE USING AVERAGE
ROI.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),ROI = longrunROI) #creating dataframe of longrun ROI of each cryptocurrency
ROI.data
```



```{r}
#filtering data to only include the relevant dates
BTC.charts2 <- BTC.charts[BTC.charts$date >= '2016-01-10' & BTC.charts$date <= '2021-09-01',]
DASH.charts2 <- DASH.charts[DASH.charts$date >= '2016-01-10' & DASH.charts$date <= '2021-09-01',]
ETH.charts2 <- ETH.charts[ETH.charts$date >= '2016-01-10' & ETH.charts$date <= '2021-09-01',]
LTC.charts2 <- LTC.charts[LTC.charts$date >= '2016-01-10' & LTC.charts$date <= '2021-09-01',]
PPC.charts2 <- PPC.charts[PPC.charts$date >= '2016-01-10' & PPC.charts$date <= '2021-09-01',]
XLM.charts2 <- XLM.charts[XLM.charts$date >= '2016-01-10' & XLM.charts$date <= '2021-09-01',]
XRP.charts2 <- XRP.charts[XRP.charts$date >= '2016-01-10' & XRP.charts$date <= '2021-09-01',]
```

```{r}
#using for loop and vectorization to get mean daily return for each currency and formatting it in data frame.
#we find out that ETH has the best mean daily return

#NO CHANGE IN RESPONSE NOW THAT WE ARE USING AVERAGE
chart.vector2 <- c("BTC.charts2","DASH.charts2","ETH.charts2","LTC.charts2","PPC.charts2","XLM.charts2","XRP.charts2")

i <- 1

for (j in chart.vector2){
    eval(parse(text=paste0("dailyReturn",i," <- (",j,"$avg[2:nrow(",j,")]-",j,"$avg[1:nrow(",j,")-1])/",j,"$avg[1:nrow(",j,")-1]")))
  i <- i+1
} 

meanDailyReturn.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),MeanDailyReturn=c(mean(dailyReturn1),mean(dailyReturn2),mean(dailyReturn3),mean(dailyReturn4),mean(dailyReturn5),mean(dailyReturn6),mean(dailyReturn7)))
meanDailyReturn.data
```


```{r}
#using sd function on each of the vectors of daily returns and then formatting it in data frame shows us that BTC 
#has the least standard deviation
std.data <- data.frame(Currencies = c("BTC","DASH","ETH","LTC","PPC","XLM","XRP"),standarDeviation = c(sd(dailyReturn1),sd(dailyReturn2),sd(dailyReturn3),sd(dailyReturn4),sd(dailyReturn5),sd(dailyReturn6),sd(dailyReturn7)))
std.data
```


```{r}
#PORTFOLIO 1 VALUE AS OF 2021 SEPT
numBTC=5000/BTC.charts2[1,5]
pf1Val=BTC.charts2[2062,5]*numBTC
pf1Val
```


```{r}
#PORTFOLIO 2 VALUE AS OF 2021 SEPT
numXRP=2500/XRP.charts2[1,5]

numLTC=2500/LTC.charts2[1,5]

pf2val=(LTC.charts2[2062,5]*numLTC)+(XRP.charts2[2062,5]*numXRP)
pf2val
```
```{r}
(pf1Val-5000)/5000
(pf2val-5000)/5000
(pf3val-5000)/5000
```


```{r}
#PORTFOLIO 3 VALUE AS OF 2021 SEPT
numETH=1250/ETH.charts2[1,5]
numDASH=1250/DASH.charts2[1,5]
numPPC=1250/PPC.charts2[1,5]
numXLM=1250/XLM.charts2[1,5]

pf3val=(ETH.charts2[2062,5]*numETH)+(DASH.charts2[2062,5]*numDASH)+(PPC.charts2[2062,5]*numPPC)+(XLM.charts2[2062,5]*numXLM)
pf3val

```


```{r}
#VALUE OF EACH PORTFOLIO IN OUR TIME PERIOD OF INTEREST
pf1runval <- data.frame(date=BTC.charts2$date[1:2062],runningValue= numBTC*BTC.charts2$avg[1:2062])

pf2runval <- data.frame(date=XRP.charts2$date[1:2062],runningValue = ((XRP.charts2$avg[1:2062]*numXRP)+(LTC.charts2$avg[1:2062]*numLTC)) )

pf3runval <- data.frame(date=ETH.charts2$date[1:2062], runningValue = (ETH.charts2$avg[1:2062]*numETH)+(DASH.charts2$avg[1:2062]*numDASH)+(PPC.charts2$avg[1:2062]*numPPC)+(XLM.charts2$avg[1:2062]*numXLM))

```

```{r}
#COMBINED DATAFRAME OF ALL PORTFOLIO VALUES
pfValues <- data.frame(Date = BTC.charts2$date, Portfolio1 = pf1runval$runningValue, Portfolio2 = pf2runval$runningValue, Portfolio3 = pf3runval$runningValue)
```

```{r}
head(pfValues,3)
```

```{r}
#PLOT OF ALL 3  PORTFOLIO VALUES
ggplot(data=pfValues,aes(x=Date))+geom_line(aes(y=Portfolio1,color = "Portfolio1"))+geom_line(aes(y=Portfolio2,color = "Portfolio2"))+geom_line(aes(y=Portfolio3,color = "Portfolio3"))+labs(y="Value ($)",x="Date")+ggtitle("Value of Portfolios")+theme(plot.title=element_text(hjust=0.5))+ylim(0,5200000)
```
```{r}
#THE MAXIMUM VALUES ATTAINED BY EACH PORTFOLIO QUESTION 2B
cat("The maximum dollar values of Portfolio 1,2 and 3 in order are",max(pfValues$Portfolio1),"and",max(pfValues$Portfolio2),"and",max(pfValues$Portfolio3))

```
```{r}
pf2dailyReturn <- c(NA,(pf2runval$runningValue[2:2062]-pf2runval$runningValue[1:2061])/pf2runval$runningValue[1:2061])
pf3dailyReturn <- c(NA,(pf3runval$runningValue[2:2062]-pf3runval$runningValue[1:2061])/pf3runval$runningValue[1:2061])
pf1dailyReturn <- c(NA,(pf1runval$runningValue[2:2062]-pf1runval$runningValue[1:2061])/pf1runval$runningValue[1:2061])
```

```{r}
#2C SCATTERPLOT 1
qplot(x=na.omit(pf2dailyReturn),y=dailyReturn1)+labs(x='Portfolio 2 Daily Return',y='BTC Daily Return')+ggtitle("Scatterplot of BTC Daily Returns vs Portfolio 2")
```

```{r}
#2C SCATTERPLOT 2
qplot(x=na.omit(pf3dailyReturn),y=dailyReturn1)+labs(x='Portfolio 3 Daily Return',y='BTC Daily Return')+ggtitle("Scatterplot of BTC Daily Returns vs Portfolio 3")
```

```{r}
#WE CAN SEE THERE IS NO CHANGE, CORRELATION STILL STRONGER FOR PF3 VS PF1 AS OPPOSED TO PF2 VS PF1
cor(na.omit(pf2dailyReturn),dailyReturn1)
cor(na.omit(pf3dailyReturn),dailyReturn1)
```


```{r}

cor(na.omit(pf1dailyReturn),na.omit(pf3dailyReturn))
```

```{r}
#adding daily returns to the running value dataframe
pf1runval$dailyReturnpf1 <- pf1dailyReturn
pf2runval$dailyReturnpf2 <- pf2dailyReturn
pf3runval$dailyReturnpf3 <- pf3dailyReturn
```

```{r}
#ADDING BITCOIN DATA INTO THE PORTFOLIO DATAFRAME FOR EASY ANALYSIS
pf2withBTC <- data.frame(date=pf2runval$date,Portfolio2value=pf2runval$runningValue,Portfolio2DailyReturn=pf2runval$dailyReturnpf2,BTCDailyReturn=c(NA,dailyReturn1))
head(pf2withBTC,3)
```

```{r}
#ADDING BITCOIN DATA INTO THE PORTFOLIO DATAFRAME FOR EASY ANALYSIS
pf3withBTC <- data.frame(date=pf3runval$date,Portfolio3value=pf3runval$runningValue,Portfolio3DailyReturn=pf3runval$dailyReturnpf3,BTCDailyReturn=c(NA,dailyReturn1))
head(pf3withBTC,3)
```


```{r}
#FILTERING FOR DAYS WHERE BTC HAS NEGATIVE OR POSITIVE RETURNS
pf2withBTCplus <- filter(pf2withBTC,BTCDailyReturn > 0)
pf2withBTCneg <- filter(pf2withBTC,BTCDailyReturn < 0)
pf3withBTCplus <- filter(pf3withBTC,BTCDailyReturn > 0)
pf3withBTCneg <- filter(pf3withBTC,BTCDailyReturn < 0)
```

```{r}
#FINDING MEAN RETURNS OF THE PORTFOLIO - WE CAN SEE THAT PORTFOLIO 2 DOES WORSE WHEN BTC RETURNS DROP THAN IT DOES WHEN BTC RETURNS GO UP. THIS IS DIFFERENT COMPARED TO OUR ORIGINAL RESPONSE USING CLOSING PRICE
mean(pf2withBTCplus$Portfolio2DailyReturn)
mean(pf2withBTCneg$Portfolio2DailyReturn)
mean(pf3withBTCplus$Portfolio3DailyReturn)
mean(pf3withBTCneg$Portfolio3DailyReturn)
```















